Pulmonary Exacerbations in CF

### 2024-01-02 Scott Sagel/PEx in CF/analysis_for_yaron.qmd
initial analysis for Yaron finished

```{r data cleaning}
# Import
df <- read.csv("./Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
  na.strings = ""
)
# Get admit year (some missing in admit_year field)
df$year <- year(df$admit_date)
# Labels
labels <- list(
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "NTM testing performed?" = "ntm_test",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "A. xylosoxidans" = "bacteria_6months___1",
  "B. cepacia complex" = "bacteria_6months___2",
  "H. influenzae" = "bacteria_6months___3",
  "MRSA" = "bacteria_6months___4",
  "MSSA" = "bacteria_6months___5",
  "P. aeruginosa" = "bacteria_6months___6",
  "S. maltophilia" = "bacteria_6months___7",
  "Other bacteria" = "bacteria_6months___8"
)
labels <- setNames(paste0(names(labels)," (",labels,")"), labels)
# Format
t <- df %>%
  select(
    sid, redcap_event_name, year, redcap_data_access_group,
    fungal_test,fungi___1:fungi___5,
    ntm_test, ntm___1:ntm___3,
    bacteria_6months___1:bacteria_6months___8
  ) %>%
  filter(!is.na(year))
```

```{r}
#| label: tbl-1
#| tbl-cap: Variable names
t %>%
  select(-sid, -redcap_event_name) %>%
  tbl_summary(
    by = year, missing_text = "Missing",
    label = labels
  ) %>%
  add_overall()
```

```{r include=FALSE}
# Write data to share with Yaron and Elin
write.csv(t,
  file = "./Scott Sagel/PEx in CF/Data_Cleaned/data_clean.csv",
  row.names = F, na = ""
)
```

>>>>

```{r data cleaning}
# Import
df <- read.csv("./Scott Sagel/PEx in CF/Data_Raw/ExacerbationsInCF222_DATA_2024-01-02_0857.csv",
  na.strings = ""
)
# Get admit year (some missing in admit_year field)
df$year <- year(df$admit_date)
# Rename centers
df$redcap_data_access_group <- factor(df$redcap_data_access_group,
  levels = c("chco", "childrens_national", "orange_county", "rady", "seattle"),
  labels = c("CHCO", "Children's National", "Orange County", "Rady", "Seattle")
)
df$cftr_mod <- factor(df$cftr_mod, levels = 0:1, labels = c("No", "Yes"))
# Labels
labels <- list(
  "Aspergillus fumigatus" = "fungi___1",
  "Aspergillus (other species)" = "fungi___2",
  "Scedosporium apiospermum" = "fungi___3",
  "Candidia (any species)" = "fungi___4",
  "Other fungus" = "fungi___5",
  "NTM testing performed?" = "ntm_test",
  "M. abscessus species" = "ntm___1",
  "M. avium complex species" = "ntm___2",
  "Other NTM species" = "ntm___3",
  "A. xylosoxidans" = "admit_cx_results___1",
  "B. cepacia complex" = "admit_cx_results___2",
  "H. influenzae" = "admit_cx_results___3",
  "MRSA" = "admit_cx_results___4",
  "Methicillin-susceptible Staph aureus" = "admit_cx_results___5",
  "P. aeruginosa" = "admit_cx_results___6",
  "P. aeruginosa mucoid" = "admit_cx_results___7",
  "S. maltophilia" = "admit_cx_results___8",
  "Other bacteria" = "admit_cx_results___9",
  "CFTR Modulator" = "cftr_mod",
  "Cultures Tested" = "redcap_data_access_group"
)
labels <- setNames(names(labels), labels)
# Format
df <- df %>%
  select(
    sid, redcap_event_name, redcap_data_access_group, year, cftr_mod,
    fungal_test, fungi___1:fungi___5,
    ntm_test, ntm___1:ntm___3,
    admit_cx_results___1:admit_cx_results___9
  ) %>%
  filter(!is.na(year))
```

# Fungi

The tables below are limited to patients who had fungal testing performed.

```{r}
#| label: tbl-fungi
#| tbl-cap: Fungi
df %>%
  filter(fungal_test == 1) %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-fungi_center
#| tbl-cap: Fungi by Center
#| message: false
df %>%
  filter(fungal_test == 1) %>%
  select(year, redcap_data_access_group, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-fungi_cftr_mod
#| tbl-cap: Fungi by Modulator Status
#| message: false
df %>%
  filter(fungal_test == 1) %>%
  select(year, cftr_mod, fungi___1:fungi___5) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

# NTM

The tables below are limited to patients who had NTM testing performed.

```{r}
#| label: tbl-ntm
#| tbl-cap: NTM
df %>%
  filter(ntm_test == 1) %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-ntm_center
#| tbl-cap: NTM by Center
#| message: false
df %>%
  filter(ntm_test == 1) %>%
  select(year, redcap_data_access_group, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-ntm_cftr_mod
#| tbl-cap: NTM by Modulator Status
#| message: false
df %>%
  filter(ntm_test == 1) %>%
  select(year, cftr_mod, ntm___1:ntm___3) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

# Bacteria

```{r}
#| label: tbl-bacteria
#| tbl-cap: Bacteria
df %>%
  select(
    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
  ) %>%
  tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: tbl-bacteria_center
#| tbl-cap: Bacteria by Center
#| message: false
df %>%
  select(
    year, redcap_data_access_group, admit_cx_results___1:admit_cx_results___9
  ) %>%
  tbl_strata(
    strata = redcap_data_access_group,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = year, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```

```{r}
#| label: tbl-bacteria_cftr_mod
#| tbl-cap: Bacteria by Modulator Status
#| message: false
df %>%
  select(year, cftr_mod, admit_cx_results___1:admit_cx_results___9) %>%
  tbl_strata(
    strata = year,
    .tbl_fun =
      ~ .x %>%
        tbl_summary(by = cftr_mod, missing_text = "Missing", label = labels) %>%
        add_overall() %>%
        add_p() %>%
        separate_p_footnotes()
  )
```
